---
title: "Case Study 4"
output: html_notebook
---

### Import Data and install/load relevant packages
```{r}
data = read.table("data.txt", head=T)
data = data[order(data['density']),]
library(L1pack) #Library for LAD = Least Absolute Deviations Regression Line
library(quantreg) #Library for Quantile Regression line
```

## Scenario 1: Fitting

### Use the data to fit the gain, or a transformation of gain, to density. Try sketching the least squares line on a scatter plot.

### Do the residuals indicate any problems with the fit?


```{r}
title = "Scatter plot of the Density and Gain"
y.axis = "Gain"
x.axis = expression('Density (g/cm'^3*')')
plot(data, main= title, pch=1, cex=1, col="orangered1", ylab=y.axis, xlab=x.axis) #Scatter plot of the data
fit = lm(gain~density, data) #Calculate the Least Squares Regression line for the given data
abline(fit, col="blue1") #Try to fit the Least Squares Regression line to the data
legend('topright', legend=c('Least Square Regression Line', 'Data'), col=c("blue1","orangered1"), lty=1)
```

### Since the data is looking like that of an exponential distribution, we will take the log of the gain variables to achieve a linear graph (to satisfy Least Squares Line condition). Moreover the data does not follow the line drawn across the graph.


```{r}
title1 = "Scatter plot of Density and log(Gain)"
y.axis1 = "log(Gain)"
log.data = data.frame(data$density, log(data$gain)) #Create a new data set where the we take the log of the gain variables
plot(log.data, main=title1, ylab=y.axis1, xlab=x.axis, pch=1, cex=1, col="darkorange")
fit.log = lm(log.data.gain.~data.density, log.data)
abline(fit.log, col="blue1")
legend('topright', legend=c('Least Square Regression Line', 'Data'), col=c("blue1","darkorange"), lty=1)
```

### Since there are replicated gain measurements, we will average thereplicate measurements to create a single variable that corresponds to the variables (explanatory and response variables need to be linear to meet the conditions required for Least Squares Line). Notice that the data fits to the regression line.


```{r}
title2 = "Scatter plot of the average of Density and log(Gain)"
mean.log.data = aggregate.data.frame(list(gain=log.data$log.data.gain.),list(density=log.data$data.density), FUN=mean)
plot(mean.log.data, main=title2, xlab=x.axis, ylab=y.axis1, pch=1, cex=1, col="darkgoldenrod1")
fit.mean.log = lm(gain~density, mean.log.data)
abline(fit.mean.log, col="blue1")
legend('topright', legend=c('Least Square Regression Line', 'Data'), col=c("blue1","darkorange"), lty=1)
```

### Perform least squares regression to fit the line to the data

```{r}
LS = lm(gain~density, data=mean.log.data) #Least Square regression
LAD = lad(gain~density, data=mean.log.data) #Least Absolute Deviations Regression
QRL = rq(gain~density, data=mean.log.data) #Quantile Regression Line
plot(log.data, main=title, xlab=x.axis, ylab=y.axis1)
legend("topright", legend=c('Least Squares', 'Least Absolute Deviation','50% Quantile'), col=c('green','blue','red'), lty=1)
abline(LS, col="green", lwd=2) #Draw a line for Least Square Regression
abline(LAD, col="blue", lwd=2) #Draw a line for LAD Regression
abline(QRL, col="red", lwd=1) #Draw a line for 50% Quantile Regression

plot(log.data, xlim=c(0.3,0.34), ylim=c(4.2,4.8), main = title, xlab=x.axis, ylab=y.axis1) #Zoomed in version to visualize regression lines
legend("topright", legend=c('Least Squares', 'Least Absolute Deviation','50% Quantile'), col=c('green','blue','red'), lty=1)
abline(LS, col="green", lwd=2)
abline(LAD, col="blue", lwd=5)
abline(QRL, col="red", lwd=2)  

cor = cor(mean.log.data) #Find the correlation coefficients
cor.rsq = append(cor, summary(LS)$r.squared) #A vector of the correlation coefficients and r squared value

LS 
LAD 
QRL 
cor.rsq 
```

### We will proceed further and check the conditions required for linear regression. Such conditions include Linearity, Normality of the residuals and Constant Variability.

```{r}
title3 = "Residuals of log(Gain)"
LAD.Residual.gain = log.data['log.data.gain.'] - rep(predict(LAD), each=10) #Calculate for the LAD gain
QRL.Residual.gain = log.data['log.data.gain.'] - rep(predict(QRL), each=10) #Calculate for the QRL gain

LAD.Residual = data.frame(log.data['data.density'], LAD.Residual.gain)
QRL.Residual = data.frame(log.data['data.density'], QRL.Residual.gain)

plot(fit.log$residuals, main='Residual plot of the Least Squares Regression Line', ylab=y.axis1, col='green4') #Residual plot of Least Sqaures
abline(0,0,col='red') #A fitted line to the residual plot
plot(LAD.Residual$log.data.gain., main='Residual plot of LAD Line', ylab='Gain', col='blue2') #Residual plot of LAD
abline(0,0,col='red') #A fitted line to the residual plot
plot(QRL.Residual$log.data.gain., main='Residual plt of Quantile Regression Line', ylab='Gain', col='darkmagenta') #Residual plot of Quantile Regression
abline(0,0,col='red') #A fitted line to the residual plot



hist((fit.log$residuals),xlim=c(-0.2,0.3), col='coral3', xlab=y.axis, main='Histogram: Least Squares Regression Line', breaks=15) #Histogram of the least sqaures plot is normal
qqnorm(fit.log$residuals, col='burlywood4') #qqplot of the Least Squares residual plot 
qqline(fit.log$residuals, col='red1')

hist(LAD.Residual$log.data.gain., xlim=c(-0.2,0.3), xlab='Gain', main='Histogram: LAD Regression Line', breaks=15, col='darkgoldenrod3') #Histogram of the LAD is normal
qqnorm(LAD.Residual$log.data.gain., col='darkolivegreen') #qqplot of the LAD residual plot
qqline(LAD.Residual$log.data.gain., col='red1')

hist(QRL.Residual$log.data.gain., xlim=c(-0.2,0.3), xlab='Gain', main='Histogram: LAD Regression Line', breaks=15, col='darkblue') #Histogram of the LAD is normal
qqnorm(QRL.Residual$log.data.gain., col='blueviolet') #qqplot of the QRL residual plot
qqline(QRL.Residual$log.data.gain., col='red1')



plot(fit.log$residuals, ylim=c(-.6,.6), main= 'Adjusted Residual plot of Least Squares Regression Line', ylab=y.axis1, col='green4') #Check for constant Variability with an adjusted plot for Least Squares
abline(0,0,col='red')
plot(LAD.Residual$log.data.gain., ylim=c(-.6,.6), main='Adjusted Residual plot of LAD Line', ylab='Gain', col='blue2') #Check for constant Variability with an adjusted plot for LAD
abline(0,0,col='red') #A fitted line to the residual plot
plot(QRL.Residual$log.data.gain., ylim=c(-.6,.6), main='Adjusted Residual plt of Quantile Regression Line', ylab='Gain', col='darkmagenta') #Check for constant Variability with an adjusted plot for Quantile Regression
abline(0,0,col='red') #A fitted line to the residual plot
```

## Scenario 3: Cross Validation

### To check how well your procecure works, omit the set of measurements corresponding to the block of density 0.508, apply your "estimation"/calibration procedure to the remaining data, and provide an interval estimate for the density of a block with an average reading of 38.6.
### Where does the actual density fall in the interval? Try the same test, for the set of measurements at the 0.001 density.

```{r}
omit.log.data = log.data[-c(1:10),] #omit data
omit.mean.log.data = mean.log.data[-1,] #omit data

New.LS = lm(gain~density, omit.mean.log.data) #Find the Least Squares Regression Line for the new data
avg.density = mean(omit.mean.log.data$density) #Find the mean of the density for the new data
sigma = sum((omit.mean.log.data$density - avg.density)^2) 

plot(omit.log.data, main='Scatter plot of the Density and Gain with omitted data', xlab=x.axis, ylab=y.axis1)
abline(LS, col='red1')
legend('topright', legend=c('Least Squares Regression Line'), col=c('red'), lwd=1)

s2 = aggregate(list(var=LS.Residual$LS.Residual.gain), by=list(density=LS.Residual$data.density), FUN=var)
s =sqrt(mean(s2$var))

bounds = 10*s*sqrt((1/9-1) + [(density-avg.density)^2/sigma)]

PredictLogGain <- function(density)
predict(LS, data.frame(density=density)) 
 
```
