---
title: "HT and CI for Paired Populations"
output: html_notebook
---
##### Do the older man and younger man in gay couples have similar levels of attractiveness (as judged by a committee of raters)? Load the couples.csv file. This contains the attractiveness scores for the older and younger men in 1000 gay couples (each score is the average of 15 "attractiveness raters", where 0 is less attractive and 10 is more attractive). Note: The meaning of "attractiveness" is complex and varies from person to person, which is one reason psychology researchers might use the avg rating from some fixed committee of people. In this way, you don't need to ever define "attractiveness", it is just the avg of all the individual definitions of the people on your commitee (whatever that means!). Read in the data set and conduct a hypothesis test with $\alpha=0.04$ to answer this question. Afterward, make a CI that has duality with this particular hypothesis test. You should use R commands to do these tasks (you can check the R commands using pencil and paper approaches if you wish). Include your code and important findings.


##### Part A: Conduct a hypothesis test with $\alpha=0.04$.

Note the formula for t is the following:
$T=\frac{\bar D-0}{{s_d}/{\sqrt n}}$
```{r}
data = read.csv('couples.csv', head=T) #load the dataset

a = 0.04 #significance level
n = 1000 #sample size
u_o = mean(data$olderman) #mean of the attractiveness of the olderman
u_y = mean(data$youngerman) #mean of the attractiveness of the youngerman
s_o = sd(data$olderman) #sd of the attractiveness of the olderman
s_y = sd(data$youngerman) #sd of the attractiveness of the youngerman
s_d = s_y - s_o

df = n - 2
#sp2 = ((n-1)*s_o^2+(n-1)*s_y^2)/(n+n-2) #calculate for sp^2
#sp = sqxrt(sp2) #don't forget to take the sqrt of sp^2
t = (u_y - u_o) / (s_d/sqrt(n)) #calculate for the t-value
p = 2*pt(t, df=df, lower=F) #calcuate of the p-value

print(p)
p < a #Check if p-value if less than a = 0.04
cat('The p-value is', p, 'so p < a is', p < a, 'hence we reject the null hypothesis!')
```
Since we have that p < a is true, we reject the null hypothesis, which means that older man and younger man in gay couples do not have similar levels of attractiveness!



##### Part B: Make a CI that has duality.

Recall that the formula for a CI is the following:
CI:$(\bar D - t_{\frac{\alpha}{2},n+m-2}\frac{s_d}{\sqrt{n}},\bar D + t_{\frac{\alpha}{2},n+m-2}\frac{s_d}{\sqrt{n}})$
```{r}
dbar = u_y - u_o #calculate for xbar, which is the difference between the two means
t_a2 = qt(0.02, df=df, lower=F) #value for t_alpha/2

lb = dbar - t_a2*(s_d/sqrt(n))
ub = dbar + t_a2*(s_d/sqrt(n))
cat('The CI is :','(',lb,ub,')')

```
We can check our answer using the 't-test function
```{r}
diff = data$younger - data$olderman
t.test(diff, conf.level = 1-a)
```



