---
title: "MATH 185 - Computational Statistics - Homework 1"
author: Dukki Hong
PID: A98058412
Section: A04
geometry: margin=2cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
---
## Problem 1

We want to compute the power of the chi-squared test by simulation. Here is a simple setting. Suppose that the null distribution (denoted $P_0$) is the uniform distribution on $\{1,...,2k\}$, where $k\geq 1$ is an integer. We want to compute the power of the chi-squared test against $P_t$, which has same support, and defined as $P_t(j)=\frac{1}{2k} +t$ for $j\leq k$ and $P_t(j)=\frac{1}{2k} -t$ for $j\geq k+1$. We assume that $t\in(0,\frac{1}{2k})$.

### Part A

Write an $R$ function chisq.power$(k,t,n,B = 2000)$ that does the following. The function creates an empty binary vector $R$ of length $B$. For each $b = 1,...,B$, the function first simulateds $X_1^b,...,X_n^b$ iid from $P_t$ (with the specified $k$) and then sets $R_b = 1$ if the chi-squared test calibrated at the $5%$ level rejects, and $R_b=0$ otherwise. When the loop over $b$ ends, the function returns the proportion of entries in $R$ that are equal to $1$. (this estimates the power of the test. Can yous ee this?)

```{r}
chisq.power <- function(k,t,n,B = 2000){
  R <- vector(mode='numeric', length = B)     #create an empty binary vector of length B
  
  null <- c(1:(2*k))                          #data under the null, uniformly distributed
  
  obs <- c()                               #create a list
  for (i in 1:(2*k)){                         #Generate data for Observed counts
    j <- c(1:(2*k))
  
    if (j[i] <= k){
      P_t <- 1/(2*j[i]) + t
    }
    else {
      P_t <- 1/(2*j[i]) - t
    }
  
    obs <- c(obs, P_t)
  }
  
  sim = sample(obs, n, replace = T, prob = rep(1/(2*k), times=(2*k))) 
  
  for(i in 1:(2*k)){
    if (chisq.test(null, obs)$p.value <= 0.05){
      replace(R, i, 1)
    }
    print(R)
  }
}
```


### Part B

## Problem 2

Are the chance of a baby being born a girl the same across counties in California? Go
to http://wonder.cdc.gov/natality.html. In Section 1, choose Gender and County. In Section
2, choose the state to be California. In Section 4, choose the year 2017. Then click on \Send"
anywhere. Enter the data in R. You can do so by hand or click on \Export", edit the resulting text
le and then use the function read.table. (You will have to edit the TXT le in order to read it into
R.) Save the dataset as an RDA le named natality-california-2017.rda. Do this as a preprocessing.
Start your code by loading the dataset using the function load.

Use these data to answer the question the best you can. Start by formalizing the question into a hypothesis testing problem, display relevant summary statistics and graphics, and then perform an appropriate test. Conclude with a sentence or two.

<font color=red>$H_0:(p_1,...,p_n)=(p_1^0,...,p_1^0)$, Chance of baby being born a girl SAME across counties in CA</font>

<font color=red>$H_1:(p_1,...,p_n)\neq (p_1^0,...,p_1^0)$, Chance of baby being born a girl DIFFERENT across counties in CA</font>
```{r}
mydata <- read.table("~/Desktop/UCSD MATH/Math.185/Natality, 2007-2017.txt", header = TRUE)
save(mydata, file="~/Desktop/UCSD MATH/Math.185/natality-california-2017.rda")

load("natality-california-2017.rda") 

mydata = data.frame(mydata) #change mydata into a dataframe
mydata = subset(mydata, select=c('Gender', 'County.Code','Births')) #remove unnecessary columns

#Are the chance of a baby being born a girl the same across counties in California?
#Parameter of interest: chance of baby being born a girl (p_i)
#H_0: chance of baby being born a girl SAME across counties in CA
#H_1: chance of baby being born a girl DIFFERENT across counties in CA

mydata = mydata[order(mydata$County.Code),] #sort out 'County.Code' column in an increasing order

female = mydata[seq(1,72,2),] #sort out mydata with females only
male = mydata[seq(2,72,2),] #sort out mydata with males only

f = matrix(female$Births, nrow = 1, ncol=36) #turn data into matrices
m = matrix(male$Births, nrow = 1, ncol=36)

X = rbind(f,m) #combine the two matrices into one matrix
row.names(X) = c('Female','Male') #Label row names 
  
chisq.test(X)

cat('value of X-Squared is', '41.285,', 'df is 35', 'and p-value is:', '0.215.')
```

<font color=red>At $\alpha =0.05$ level of significance, we have the p-value $p=0.215$ ($p=0.215 > 0.05 = \alpha$). Hence we keep the null hypothesis ($H_0$). This means that the chance of baby being born a girl is same across counties in California.</font>

*Now we proceed to showing a graphical representation of the data*
```{r}
col = c('pink','skyblue') #set color
ylab = ('Number of births') #Y-label name
barplot(X, legend=T, col=col, ylab=ylab, ylim=c(0,120000)) 
barplot(X, , beside= T, legend=T, col=col, ylab=ylab)
```


## Problem 3

Write a function *chisq.perm.test*(tab;B = 2000) implementing the chi-squared test of independence calibrated by permutation. The inputs are a table of joint counts (without totals) and number of permutations to be done. Have the test return the p-value. Test your function on the *HairEyeColor* dataset.

```{r}
chisq.perm.test <- function(tab, B=2000){
  obs = chisq.test(tab)$stat #observed statistic from the Chisqd test
  
  nrow = nrow(tab) #Number of rows
  ncol = ncol(tab) #Number of columns
  
  rows = numeric(nrow) #create a 0 vector with n number of rows
  totrow = numeric(0) #create a vector with 0's
  
  for (i in 1:nrow){ #for-loop to fill up the vectors defined above
    rows[i] = sum(tab[i,]) #summing up each of the rows
    totrow = c(totrow, rep.int(i, times=rows[i]))
  }
  
  #do the same with the columns
  cols = numeric(ncol) #empty vector with n columns
  totcol = numeric(0) #0 vector
  
  for (i in 1:ncol){
    cols[i] = sum(tab[,i]) #summing up each of the columns
    totcol = c(totcol, rep.int(i, times=cols[i]))
  }
  
  D = numeric(0) #vector for permutations for the Chisqd test statistic
  tot = sum(tab[,]) #total number of counts
  
  #####################################
  
  for (b in 1:B){
    x = totrow
    y = sample(totcol, tot, replace=FALSE) #sample column totals 'tot' times without replacement
    matperm = cbind(x,y) #combine columns of x and y
    
    matpermcount = matrix(0, nrow=nrow, ncol=ncol) #create a 0 matrix with n number of rows and columns
    
    for(x in 1:nrow){ #rows from 1 - 4
      for(y in 1:ncol){ #rows from 1 - 4
        for(m in 1:tot){ #loop for going through all the rows in matperm
          if(all(matperm[m,] == c(x,y)) == TRUE){ #ordered pairs such as (1,1), (2,2) and etc...
            matpermcount[x,y] = matpermcount[x,y] + 1
          }
        }
      }
    }
  
    D[b] = chisq.test(matpermcount)$stat #chisq.test statistic for a given permutation
  
    #Check if chisq.test stat >= observed stat
    count = 0 #counts to check (chisq.test stat > observed)
    if (D[b] >= obs){
      count = count + 1
    }
    return((count+1)/(B+1))
  }
}
```

<font color=red>Now we check if the function works<font>
```{r}
tab = apply(HairEyeColor, c(1,2), sum)
tab
chisq.perm.test(tab)
cat('We get that the Chi-squared test statistic is equal to :',chisq.perm.test(tab))
```



## Problem 4

Go to the following webpage: https://catalog.data.gov/dataset/school-improvement-2010-grants 
Download the dataset. The description is on the webpage. To read it into R, use the function
read.csv. Remove the schools from Rhode Island as their selected models are missing. Do this on your own to practice reading datasets into R. Save the dataset as an RDA le named school-improvement-2010.rda. Do this as a preprocessing. Start your code by loading the dataset using the function load.

We are interested in the following question: Is there an association between the model that each school selected and the state where the school was located at that time?

### Part A

Explore this question with one or several appropriate plots. Then formulate the question into a hypothesis testing problem and perform a test. Conclude with some brief comments.

### Part B

In this particular case, is the method of Problem 3 applicable? If so, apply it and compare with the previous test that you performed.


